---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%" # ,
  # eval = selenider::selenider_available()
)
```

# selenider

<!-- badges: start -->
[![R-CMD-check](https://github.com/ashbythorpe/selenider/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ashbythorpe/selenider/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/ashbythorpe/selenider/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ashbythorpe/selenider?branch=main)
<!-- badges: end -->

selenider aims to make web testing and scraping in R much simpler, using either [chromote](https://rstudio.github.io/chromote/) or [RSelenium](https://www.selenium.dev/).
It is inspired by Java's [Selenide](https://selenide.org/) and Python's [Selene](https://yashaka.github.io/selene/).

It provides the following features:

* All elements are lazy, meaning that they will only actually collect an element when absolutely necessary.
* Automatic implicit waiting, meaning that functions will automatically wait a certain amount of time (by default, 4 seconds) before performing some action on it.
* An API that is designed for use with the pipe operator (`|>` or `%>%`).
* A quick and concise way to specify test expectations.

## Installation

You can install the development version of selenider like so:

``` r
# install.packages("remotes")
remotes::install_github("ashbythorpe/selenider")
```

## Usage

```{r}
library(selenider)
```

To open a session, and navigate to a website, use `open_url()`.

``` {r}
open_url("https://www.r-project.org/")
```

Use `s()` to select an element. By default, CSS selectors are used, but other options are available.

``` {r}
r_header <- s("#r-project")

r_header

html_text(r_header)
```

For example, an xpath can be used instead.

```{r}
s(xpath = "//div/a")
```

Use `ss()` to select multiple elements
```{r}
all_links <- ss("a")

all_links

html_attr(all_links[[1]], "href")
```

Use `html_element()` and `html_elements()` to find child elements of an existing element. These can be
chained with the pipe operator (`|>`) to specify paths to elements.

``` {r}
s(".row") |>
  html_element("div") |>
  html_element("p") |>
  html_text()
```

Use `html_children()` and friends to find elements using their relative position to another.
```{r}
s("p") |>
  html_children()

s("p") |>
  html_ancestors()
```

You can use `html_filter()` and `html_find()` to filter collections of elements using a custom function.
`html_find()` returns the first matching element, while `html_filter()` returns all matching elements.

``` {r}
all_links <- ss("a")

ss("a") |>
  html_find(has_text("About R")) |>
  html_attr("href")

ss("p") |>
  html_filter(is_visible)
```

selenider elements are *lazy*, meaning that when you specify the path to an element or group of
elements, they are not actually located in the DOM until you *do* something with them.

There are three types of functions that force an element to be collected:
* properties (e.g. `html_text()`)
* conditions (e.g. `is_visible()`)
* actions (e.g. `click()`)

selenider provides a concise testing interface using the `html_expect()` function. Provide an element,
and one or more conditions, and the function will wait until all the conditions are met. Conditions
can be functions or simple calls (e.g. `has_text("text")`)

``` {r}
s("#getting-started") |>
  html_expect(is_present) |>
  html_expect(has_text("Getting Started"))

s(".row") |>
  html_element("#r-foundation") |>
  html_expect(is_visible, is_enabled)

s("div[role='navigation']") |>
  html_element(".row") |>
  html_children() |>
  html_expect(has_at_least(2))

s("#help-with-r") |>
  html_expect(
    \(x) substring(html_text(x), 1, 1) == "H"
  )
```

Errors try to give as much information as possible.
```{r, error = TRUE}
s("#news") |>
  html_expect(has_text("Getting Started"))
```

And (`&&`), or (`||`) and not (`!`) can be used as if the conditions were logical values. Additionally,
you can omit the first argument to `html_expect()` (but all conditions must be calls).

``` {r}
s(".random-class") |>
  html_expect(!is_present)

s(".mt-timeline") |>
  html_expect(is_visible || is_enabled)

elem_1 <- s(".random-class")

elem_2 <- s("#getting-started")

# Test that either the first or second element exists
html_expect(is_present(elem_1) || is_present(elem_2))
```

<!-- TODO: Link to vignettes -->
